---
title: "Flow chart information: count missing."
output:
  pdf_document:
    includes:
      in_header: ../preamble-latex-nobookdown.tex
    number_sections: yes
    toc: no
    highlight: tango
  html_document:
    theme: united
    toc: no
editor_options:
  chunk_output_type: console
header-includes:
 \usepackage{float}
---

# Section 2, premenopause # {#s2pre}


<!-- Source: U:\projects\Iron-nails-serum\sections\section6.Rmd  -->


```{r, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      results = 'hide',
                      warning = F,
                      message = F,
                      error = F,
                      fig.pos = 'H')

#runif(1, 0, 10^8)
set.seed(74268794)

```


```{r, include=F}
# bring in packages
require(knitr)
require(haven)
require(data.table)
require(ggplot2)
require(kableExtra)
library(dplyr)
library(tidyr)
library(plyr)
#library(expss) # https://cran.r-project.org/web/packages/expss/vignettes/labels-support.html
# NOTE: expss seems to mask a a lot of functions from other packages listed here.
library(cowplot)
library(splines)
library(sjlabelled)
require(dplyr)

```




```{r}

# source: section2-dataprep.Rmd

load(file="../sections/s2.RData") # s2.dat, cov.vars.new, labels.df from section2-dataprep.Rmd

names(s2.dat)
dim(s2.dat)
cov.vars.new

```

```{r, eval=F, include=F}

cov.vars.new = cov.vars.new[!(cov.vars.new %in% c("HR57.2"))]; cov.vars.new

dim(s2.dat)
table(s2.dat$menop.status, useNA = "always")
s2.dat = s2.dat[s2.dat$menop.status==0,] # select premenopausal
dim(s2.dat)

s2.dat$baseline.age = scale(s2.dat$baseline.age, scale=F)
cov.vars.new

cov.vars.new = cov.vars.new[!(cov.vars.new %in% c("HR_Menopause.f",
                                                  "HR_DiffLMPintv",
                                                  "HR_HRTcurrent.f",
                                                  "rls",
                                                  "HR_HRT_Years"))]; cov.vars.new # remove menopause variable and any post-menopause related variables


table(s2.dat$menop.status, useNA = "always") # menopause variable
sum(table(s2.dat$menop.status, useNA = "always")) # menopause variable

```

# Premenopausal group

```{r}

# Missing for premenopausal group
# =================================

iron.vars = c("UMN_Iron_Baseline_FE", "UMN_Iron_Baseline_FERTN", "UMN_Iron_Baseline_FESAT")

# See groups of variables in section2-post.Rmd

# Group 1. supplements and dietary intake
# ==========================

supp.covs = c("na_all_sd_fe",
              'na_all_sd_ca_100') 

# Group 2. Lifestyle: exercise, BMI, meat consumption, alcohol consumption
# ==========================

lifestyle.covs = c("AL_DrinksPWC_4",
                 "PH2_TotHrsPerWeek_4",
                 "FG_all_pf_meat_8", 
                 'bmi')

# Group 3: years since last menstrual period, reproductive life span (years) that is calculated by subtracting age at menarche from age at menopause based on last menstrual period, estrogen or progesterone use (number of total years), and total years pregnant and breastfeeding
# ==========================

repro.covs = c( #"HR_DiffLMPintv",
                #"HR_HRT_Years",
                'totpreg.mo2')#,
                #'rls')
# take out postmenopause-related variables

# Group 4: self-report of ever having a health condition at baseline were: preeclampsia, polycystic ovary syndrome (PCOS), irritable bowel syndrome (IBS), colon/rectum polyps, and iron-deficiency anemia (IDA)
# ==========================

health.covs = c('pcos.rev', 
                'preeclampsia',
                'ibd.rev',
                'polyps.rev',
                'MC116.f')

# Group 5: : statin use, regular aspirin use (1+/week for 3 or more months), hormone replacement therapy (HRT), and recent blood donation (within past 12 months). We also included three genetic variants that have been confirmed to be associated with serum iron (rs1800562 (HFE gene), rs1799945 (HFE gene), and rs855791 (TMPRSS6 gene))
# ==========================

bin.covs = c('HZ_RX_Chol_Statin_Current',
             'HZ_RX_NSAID_Aspirin_Current_Reg',
             'blood.don.12mos')#,
             #'rs1799945.g',
             #'rs1800562.a',
             #'rs855791.g')
# genetic variables are not from self-report questions so take out 


# analysis vars
avars = c(supp.covs, lifestyle.covs, repro.covs, health.covs, bin.covs); avars

# subset to premenopausal
dat1 = s2.dat[which(s2.dat$menop.status==0), avars]
length(avars) #check
dim(dat1) # check

```

## Number of missing per variable

```{r, results='markup'}
# Source: https://stackoverflow.com/questions/24027605/determine-the-number-of-na-values-in-a-column
na_count <-sapply(dat1, function(y) sum(length(which(is.na(y)))))
kable(data.frame(100*round(na_count/nrow(dat1),3)), booktabs=T, col.names = ("Percent missing"))

```


## Number of missing per participant

```{r, results='markup'}

# source: https://stackoverflow.com/questions/37801338/count-nas-per-row-in-dataframe
miss.row = rowSums(is.na(dat1))

#table(miss.row)
#sum(table(miss.row))

no.miss.row = length(avars) - rowSums(!(is.na(dat1)))

kable(table(no.miss.row), booktabs=T)

#kable(sum(table(no.miss.row)), booktabs=T)

kable(round(100*prop.table(table(no.miss.row)),1), booktabs=T, col.names = c("No. missing", "Proportion missing"))

#table(c(miss.row, no.miss.row)) # counts
#100*round(prop.table(table(c(no.miss.row, miss.row))),2) # proportion

```

## Total number missing


```{r}

miss.pre = dim(dat1[!(complete.cases(dat1)),])[1]
notmiss.pre = dim(dat1[complete.cases(dat1),])[1]

prop.miss.pre = dim(dat1[!(complete.cases(dat1)),]) / (dim(dat1[complete.cases(dat1),]) + dim(dat1[!(complete.cases(dat1)),]))

```

`r miss.pre` missing at least one, `r notmiss.pre` missing none, and `r prop.miss.pre[1]` is the proportion missing at least one.

# Postmenopausal group

```{r}
# determine how many have just 1, 2, 3, etc... number of missing covariates.


iron.vars = c("UMN_Iron_Baseline_FE", "UMN_Iron_Baseline_FERTN", "UMN_Iron_Baseline_FESAT")

# See groups of variables in section2-post.Rmd

# Group 1. supplements and dietary intake
# ==========================

supp.covs = c("na_all_sd_fe",
              'na_all_sd_ca_100') 

# Group 2. Lifestyle: exercise, BMI, meat consumption, alcohol consumption
# ==========================

lifestyle.covs = c("AL_DrinksPWC_4",
                 "PH2_TotHrsPerWeek_4",
                 "FG_all_pf_meat_8", 
                 'bmi')

# Group 3: years since last menstrual period, reproductive life span (years) that is calculated by subtracting age at menarche from age at menopause based on last menstrual period, estrogen or progesterone use (number of total years), and total years pregnant and breastfeeding
# ==========================

repro.covs = c( "HR_DiffLMPintv",
                "HR_HRT_Years",
                'totpreg.mo2',
                'rls')


# Group 4: self-report of ever having a health condition at baseline were: preeclampsia, polycystic ovary syndrome (PCOS), irritable bowel syndrome (IBS), colon/rectum polyps, and iron-deficiency anemia (IDA)
# ==========================

health.covs = c('pcos.rev', 
                'preeclampsia',
                'ibd.rev',
                'polyps.rev',
                'MC116.f')

# Group 5: : statin use, regular aspirin use (1+/week for 3 or more months), hormone replacement therapy (HRT), and recent blood donation (within past 12 months). We also included three genetic variants that have been confirmed to be associated with serum iron (rs1800562 (HFE gene), rs1799945 (HFE gene), and rs855791 (TMPRSS6 gene))
# ==========================

bin.covs = c('HZ_RX_Chol_Statin_Current',
             'HZ_RX_NSAID_Aspirin_Current_Reg',
             'blood.don.12mos')#,
             #'rs1799945.g',
             #'rs1800562.a',
             #'rs855791.g')
# genetic variables are not from self-report questions so take out 


# analysis vars
avars = c(supp.covs, lifestyle.covs, repro.covs, health.covs, bin.covs); avars
extra.vars = c('menop.age', 'age.menarche', 'PG_BreastTotal', 'PG_MedParity') # part of rls variable

# what variables are already in the s2.dat data set?
avars[avars %in% names(s2.dat)]
extra.vars[extra.vars %in% names(s2.dat)] # none

# what analysis variables are not in the s2.dat data set?
avars[!(avars %in% names(s2.dat))]

# subset to postmenopausal
dat1 = s2.dat[which(s2.dat$menop.status==1), avars]
length(avars) #check
dim(dat1) # check

```


## Number of missing per variable

```{r, results='markup'}

# Source: https://stackoverflow.com/questions/24027605/determine-the-number-of-na-values-in-a-column
na_count <-sapply(dat1, function(y) sum(length(which(is.na(y)))))
kable(data.frame(100*round(na_count/nrow(dat1),3)), booktabs=T, col.names = c("Percent missing"))

```


## Number of missing per participant

```{r, results='markup'}
# source: https://stackoverflow.com/questions/37801338/count-nas-per-row-in-dataframe
miss.row = rowSums(is.na(dat1))

#table(miss.row)
#sum(table(miss.row))

no.miss.row = length(avars) - rowSums(!(is.na(dat1)))
kable(table(no.miss.row), booktabs=T)
#kable(sum(table(no.miss.row)), booktabs=T)
kable(round(100*prop.table(table(no.miss.row)),1), booktabs=T, col.names = c("No. missing", "Proportion missing"))

#table(c(miss.row, no.miss.row)) # counts
#100*round(prop.table(table(c(no.miss.row, miss.row))),2) # proportion

```

## Total number missing


```{r}

miss.post = dim(dat1[!(complete.cases(dat1)),])[1]
notmiss.post = dim(dat1[complete.cases(dat1),])[1]

prop.miss.post = dim(dat1[!(complete.cases(dat1)),]) / (dim(dat1[complete.cases(dat1),]) + dim(dat1[!(complete.cases(dat1)),]))

```

`r miss.post` missing at least one, `r notmiss.post` missing none, and `r prop.miss.post[1]` is the proportion missing at least one.

